---
title: "MATH1324 Assignment 2"
subtitle: "Supermarket Price Wars"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Executive Statement
The objective of the investigation is to figure out which supermarket, **Coles** or **Woolworths**, is cheaper. The sample is gathered from the website <https://grocerycop.com.au/products> which includes 9 products from each of the 10 categories. A large sample of **90 (n > 30)** is chosen in accordance with **Central Limit Theorem(CLT)** to effectively avoid the issue with normality and to **limit standard error**. We have used Stratified Sampling method to randomly select the products from each category(i.e, strata). The dataset consists of 6 variables - `Sl_No`, `Product_Name`, `Units`, `Category`, `Coles_Price`, and `Woolworths_Price`. The `Product_Name`, `Units` and `Category` match between Coles and Woolworths. All the product prices are in Australian Dollars (AUD). The summary statistics and Box plot help in comparing the prices between the stores. The QQ-Plot of `Diff` column(`Coles_Price` - `Woolworths_Price`) is used for exploring the Normality. The **paired-samples t-test** is used to check for the statistically significant mean difference between Coles and Woolworths prices. The result of the dependent sample t-test signifies that there is a **statistically significant mean difference** between Coles Price and Woolworths Price. In conclusion, **Woolworths prices are found to be significantly cheaper when compared to Coles prices.**

## Load Packages and Data
```{r}
#Loading the necessary packages and reading the dataset using read_csv() function.
library("readr")
library("magrittr")
library("car")
library("granova")
library("dplyr")
coles_vs_woolworths <- read_csv("ColesVsWoolworths.csv")
#head() function is used to display the first 10 observations of the dataset.
head(coles_vs_woolworths, 10)
#checking for any missing values in the dataset
sum(is.na(coles_vs_woolworths))
#To Display the dimensions of the dataset
dim(coles_vs_woolworths)
```

## Summary Statistics
* The mean values of `Coles_Price` and `Woolworths_Price`, obtained from the descriptive statistics, are 8.576778 and 8.097667 respectively. This shows that the average price of products in Coles supermarket is more than that of Woolworths supermarket.     
* The *Figure - 1* shows the visualisation of `Coles_Price` and `Woolworths_Price` are drawn using **Boxplot()** function.  
* The new column called `Diff` is created to store Price Difference ( `Coles_Price` - `Woolworths_Price` ).  
* The mean value of `Diff` column ( mean difference ) from descriptive statistics is found to be **0.4791111**. This shows that on average coles prices are 0.479 AUD more than Woolworths price.  
* The QQ-Plot as shown in *Figure - 2* is drawn to check for normality of the price differences. From the QQ-Plot we can see that few points are falling outside the tails of the distribution. This shows that the tails are heavier than what we would expect under a normal distribution. But, due to the large sample size( **n=90**), according to Central Limit Theorem, we can assume that the `Diff` column is normally distributed. Hence, we can proceed with the paired sample t-test.

```{r}
#Summary Statistics for Coles Prices
coles_vs_woolworths %>%
  summarise(
  Store = "Coles",
  Min = min(Coles_Price, na.rm = TRUE),
  Q1 = quantile(Coles_Price, probs = .25, na.rm = TRUE),
  Median = median(Coles_Price, na.rm = TRUE),
  Q3 = quantile(Coles_Price, probs = .75, na.rm = TRUE),
  Max = max(Coles_Price, na.rm = TRUE),
  Mean = mean(Coles_Price, na.rm = TRUE),
  SD = sd(Coles_Price, na.rm = TRUE),
  N = n(),
  Missing = sum(is.na(Coles_Price))
  )
#Summary Statistics for Woolworths Prices
coles_vs_woolworths %>%
  summarise(
  Store = "Woolworths",
  Min = min(Woolworths_Price, na.rm = TRUE),
  Q1 = quantile(Woolworths_Price, probs = .25, na.rm = TRUE),
  Median = median(Woolworths_Price, na.rm = TRUE),
  Q3 = quantile(Woolworths_Price, probs = .75, na.rm = TRUE),
  Max = max(Woolworths_Price, na.rm = TRUE),
  Mean = mean(Woolworths_Price, na.rm = TRUE),
  SD = sd(Woolworths_Price, na.rm = TRUE),
  N = n(),
  Missing = sum(is.na(Woolworths_Price))
  )
#Visualizing the Coles and Woolworths Prices using boxplot() function
boxplot(
  coles_vs_woolworths$Coles_Price,
  coles_vs_woolworths$Woolworths_Price,
  ylab = "Prices",
  xlab = "Supermarkets" , col = c("firebrick" , "darkolivegreen1") , las = 1 , main = "Boxplot of Prices in AUD of Coles and Woolworths"
  )
axis(1, at = 1:2, labels = c("Coles", "Woolworths"))
#Creating a new column called "Diff" to store Price Difference
coles_vs_woolworths <- coles_vs_woolworths %>% mutate("Diff" = Coles_Price - Woolworths_Price)
#To display the Product Name, Coles price , Woolworths Price and Diff column
head(coles_vs_woolworths[, c(2,5,6,7)])
#Summary Statistics for Price Difference(Coles_Price - Woolworths_Price).  
coles_vs_woolworths %>%
  summarise(
  Min = min(Diff, na.rm = TRUE),
  Q1 = quantile(Diff, probs = .25, na.rm = TRUE),
  Median = median(Diff, na.rm = TRUE),
  Q3 = quantile(Diff, probs = .75, na.rm = TRUE),
  Max = max(Diff, na.rm = TRUE),
  Mean = mean(Diff, na.rm = TRUE),
  SD = sd(Diff, na.rm = TRUE),
  N = n(),
  Missing = sum(is.na(Diff))
  )
#Drawing QQ-Plot for the "Diff" column 
qqPlot(coles_vs_woolworths$Diff, dist="norm" , ylab = "Price Difference")
```
<center><h4><b><u>Figure - 2</u></b></h4></center>

## Hypothesis Test
* The **paired sample t-test** is chosen because the same product is being measured in both the supermarkets.
* The paired samples t-test is used to determine whether the mean difference between Coles prices and Woolworths prices is zero. 
* For the paired samples t-test we are assuming a two-tailed test with significance level(&#945;) = 0.05.Thus the confidence interval(CI) is 95%.  
* The statistical hypothesis for the paired samples t-test are as follows:  
<center>***H~0~ : μ~Δ~ = 0 ***</center>    
<center>***H~A~ : μ~Δ~ ≠ 0***</center>
&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;where, H~0~ : there is no price difference between Coles and Woolworths.  
       &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; H~A~ : there is a price difference between Coles and Woolworths.

* The paired samples t-test is calculated using the **t.test()** function and specifying **paired = TRUE**.
* The critical value t* for the paired-sample t-test are ±1.986979.The t = 2.7221 is more extreme than 1.986979.
* **granova.ds()** is used to visualise the mean difference between `Coles_Price` and `Woolworths_Price` using a scatter plot as shown in *Figure - 3*.

```{r}
#Calculating the paired sample t-test using t.test() function
t.test(coles_vs_woolworths$Coles_Price, coles_vs_woolworths$Woolworths_Price,
       paired = TRUE,
       alternative = "two.sided", mu = 0)

#The critical value t* for the paired-sample t-test, assuming a two-tailed test with significance level = 0.05
qt(p = 0.025, df = 89)

#To visualise mean difference using a scatter plot.
granova.ds(
  data.frame(coles_vs_woolworths$Coles_Price, coles_vs_woolworths$Woolworths_Price),
  xlab = "Coles Price",
  ylab = "Woolworths Price" 
  )
```
<center><h4><b><u>Figure - 3</u></b></h4></center>
## Interpretation

A paired samples t-test has been used to test for a significant mean difference between Coles price and Woolworths price. The mean difference was found to be 0.4791111 ($SD$ = 1.670). While the price differences exhibited evidence of non-normality upon inspection of the normal Q-Q plot, the central limit theorem ensured that the t-test could be applied due to the large sample size.The values of paired samples t-test are as follows - <br> $t$($df$=89)=2.7221, $p$< 0.007804, 95% $CI$ [0.1293857 0.8288365]. <br>
As the $p$-value is less than significance level(&#945; = 0.05) and 95% $CI$ does not capture H~0~, therefore we reject H~0~. Thus there was a statistically significant mean difference between Coles and Woolworths prices.Woolworths prices were found to be significantly cheaper when compared to Coles prices.  

## Discussion 

From the interpretation, we can infer that, a paired t-test found a statistically significant mean difference between the prices of products of Coles and Woolworths.As a result, we can conclude that Woolworths is cheaper than coles.  

**Strengths of the Investigation - **  

* The investigation ensures the same Product Names, brand and its Units used at Coles and Woolworths supermarkets.    
* The dataset is tidy, easy to understand and products are differentiated into multiple categories.  

**Limitations of the Investigation - **  

* The sample size of the dataset can be increased by including all the items and categories that were not covered during this investigation.
* The products that had discounted prices were not part of the dataset.
* Woolworths and Coles have their own products which are not a part of this investigation. 


**Improvements - **  

Improvements that can be done to the investigation are :
<ul>
<li> By increasing the sample size  </li>
<li>By including items from all the categories and discounted products.</li>
<li>Woolworths and Coles have their own products which can be included in the future.</li>
<li>Customer satisfaction and product star ratings can be considered in future.</li>

</ul>










